\title{Urschleim in Silicon:\\Return Oriented Program\\Evolution with ROPER}
\author{Olivia Lucca Fraser}

\degree{Master of Computer Science}
\degreeinitial{M.C.Sc.}
\mcs
\faculty{Computer Science}
\dept{Faculty of Computer Science}
\supervisor{Nur Zincr-Heywood}
\supervisor{Malcolm Heywood}
\defencemonth{December}
\defenceyear{2018}

\dedicate{This thesis is dedicated to my children, Kai, Nahni, Sophie, Quin, and Faro, and to Andrea Shepard.}
\frontmatter
% \listofalgorithms

\begin{abstract}
  This thesis lays out the design and implementation of a system for
  evolving populations of return-oriented programs, and then proceeds to study their
  behaviour in a series of experiments. 
\end{abstract}

\begin{acknowledgements}
I would like to thank my supervisors, Nur Zincir-Heywood and Malcolm Heywood,
whose support, encouragement, and expertise have been indispensable to me. My
thanks also go to to John Jacobs at Raytheon SAS, who saw enough interest and
potential in this project to grant it his agency's financial backing, when it
was still in its early stages. And thanks to all the friends and colleagues who
helped the ideas presented in these pages come to life in conversation. Without
any hope of being exhaustive, these include Andrea Shepard, Meredith Patterson,
Pete Wolfendale, Giancarlo Sandoval, Neha Spellfish, Pompolic, Petra Kendall,
Peli Grietzer, Alice Maz, 0xdeadbabe, Reza Negarestani, Dominic Fox and others
in the Greytribe and Special Circumstances chat, Sig Cox at LambdaConf, Chris
Watts, Corrie Watts, Raphael Bronfman-Nadas, Deepthi Rajashekar, Aimee Burrows,
Rob Curry, Robert Smith, and Stephen Kelly in the NIMS laboratory, past and
present, Ed Prevost, Gurjeet Clair, Katie Sexton, Sam Kaplan, and Nick Miles at
Tenable (again, past and present), and Lilly Chalupowski at GoSecure. Thanks to Rob Piece at 2Keys for his
encouragement and support for this project.
Thanks to Amy Ireland,
Diann Bauer, Helen Hester, Katrina Burch, and Patricia Reed, of Laboria
Cuboniks, to Virginia Barrett of VNS Matrix, and to Angus MacGyver, for
being inexhaustible sources of inspiration. A long-overdue thanks goes
to Uwe Petersen and Valerie
Kerruish, for their support on an earlier research on logic and dialectics,
which was aborted due to external circumstances, and of which
the current project remains a descendant, however distant. 
And thank you to my
long-suffering parents Marion and Zachary Fraser, for the countless hours they
gave up, looking after my brood, so that I could finish this thesis while they're
still young. 

This research is supported by Raytheon SAS. The research is conducted
as part of the Dalhousie NIMS Lab at: \url{https://projects.cs.dal.ca/projectx/}
\end{acknowledgements}

\printnoidxglossaries
\clearpage
\addtocontents{gls}{\protect\baselineskip=0.7\protect\baselineskip}
\mainmatter